---
title: 'Time Series Analysis - Homework #1'
author: "Luke Schwenke"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(caret)
library(dplyr)
library(ggplot2)

```

# Question 1

Suppose E(X)=2, Var(X)=9, E(Y)=0, Var(Y)=4, and Corr(X,Y)=0.25

#### (a) Var(X+Y)

= Var(X) + Var(Y) + 2*Cov(X,Y)

= 9 + 4 + 2*Cov(X,Y)

= 13 + 2*(Corr(X,Y)*sqrt(Var(X)*Var(Y)))

= 13 + 2*(0.25*sqrt(9*4))

= 13 + 2*1.5

= 13 + 3

= **16 (Answer)**

#### (b) Cov(X, X+Y)

= Cov(X,X) + Cov(X,Y)

= Var(X) + Cov(X,Y)

= 9 + 1.5 (solved in part a)

= **10.5 (Answer)**

#### (c) Corr(X+Y, X-Y)

= Cov(X+Y, X-Y) / sqrt(Var(X+Y)*Var(X-Y))

Cov(X+Y, X-Y) 

= Cov(X,X) - Cov(X,Y) + Cov(Y,X) - Cov(Y,Y)

= Var(X) - Cov(X,Y) + Cov(Y,X) - Var(Y)

= 9 - 1.5 (solved in part a) + Cov(Y,X) - 4

= 9 - 1.5 + 1.5 -4

= 5

Cov(X,Y) = Cov(Y,X) (symmetric)

Var(X+Y) 
= 16 (answer from part a)

Var(X-Y)

= Var(X) + Var(Y) - 2*Cov(X,Y)

= 9 + 4 - 2*1.5

= 13 - 3

= 10

Cov(X+Y, X-Y) / sqrt(Var(X+Y)*Var(X-Y))

= 5 / sqrt(16 * 10)

= 5 / sqrt(160)

= **0.3952847 (Answer)**

# Question 2

#### If X and Y are dependent but Var(X) = Var(Y), find Cov(X+Y,X-Y)

Cov(X+Y,X-Y)

= Cov(X,X-Y) + Cov(Y,X-Y)

= Cov(X,X) - Cov(X,Y) + Cov(Y,X) - Cov(Y,Y)

= Cov(X,X) - Cov(Y,Y)

= Var(X) - Var(Y)

**As Var(X) = Var(Y), Cov(X+Y, X-Y) = 0 (Answer)**

# Question 3

### Suppose Yt = 5 + 2t + Xt where {Xt} is a zero-mean stationary series with autocovariance function Gamma-k

#### (a) Find the mean function for {Yt}

E[Yt] 

= E[5 + 2t + Xt]

= E[5] + E[2t] + E[Xt]

= 5 + 2t + 0

= **5 + 2t (Answer)**

#### (b) Find the autocovariance function for {Yt}

Yt = 5 + 2t + Xt

So, Xt = Yt - 5 - 2t

Let Yt - 5 - 2t = Zt

Then consider the covariance between Zt and Zt-k

Since {Xt} is a zero-mean stationary series, its autocovariance function is denoted as Gamma-k.
**Therefore, the autocovariance function for {Yt} is also Gamma-k (Answer)** 

#### (c) Is {Yt} stationary? Why or why not?

* **It is not stationary because the mean value depends on time (Answer)**

# Question 4

Plot Matching

* 1 = A 
* 2 = C
* 3 = B

# Question 5

#### Simulate a normal white noise sample of size 10 with the mean 2.3 and standard deviation 1.2, and then calculate the sample mean and sample standard deviation. Are you satisfied with your computed results? How about the sample size 10,000?

```{r q5, echo=TRUE, warning=FALSE}

set.seed(55)

white <- rnorm(n=10, mean=2.3, sd=1.2)

paste("Sample 10 Mean:", round(mean(white),2))
paste("Sample 10 Standard Deviation:", round(sd(white),2))

white_large <- rnorm(n=10000, mean=2.3, sd=1.2)
paste("Sample 10,000 Mean:", round(mean(white_large),2))
paste("Sample 10,000 Standard Deviation:", round(sd(white_large),2))

```

The output above shows the small sample's mean (2.11) and standard deviation (0.99) are quite different from the true values set. As the sample size is increased dramatically to 10,000 the sample mean (2.3) and standard deviation (1.21) values approach the true values much more closely. I am more satisfied with the results of the large sample.